\title{Assignment 1: CS 215 \\ Solutions}
\author{
	Neeraj Dhake \\ 150050022
	\and
	Rohit Kumar Jena \\ 150050061
}
\date{6th August 2016}

\documentclass[11pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{ulem}
\usepackage[margin=0.5in]{geometry}
\begin{document}
\pagenumbering{gobble}
\maketitle
\newpage
\pagenumbering{arabic}

\textbf{Honor Code:} 
\begin{itemize}
\item We pledge by our honor that we will complete the assignments in a legitimate way and will not provide or recieve any unauthorized help. 
\end{itemize}
\vspace{5mm}
\textbf{Instructions to run code: }
\begin{itemize}
	\item {
		For Question 4, the corresponding files are \textit{hw1\_q4.m} (to run the file and generate plots as well as give the relative errors for both the median and the mean) and \textit{neighbours.m} (to generate the neighbours of the array at a given index .i.e. if index is 14 then it will return \textbf{z(6:22)} ). \\
		To run the program, simply type \textit{hw1\_q4} from the MATLAB command line.
	} 
	\item {
		For Question 5, the corresponding files are \textit{hw1\_q5.m} (to read the original array from the file named \textit{input\_array.txt} and compute the mean, median, and the standard deviation). Then it prompts the user to enter the new data value. It uses the \textit{UpdateMean.m}, \textit{UpdateMedian.m}, and \textit{UpdateStd.m} files to compute the new mean, new median and new standard deviation from the old values. \\
		To run the file, simply type \textit{hw1\_q5} from the MATLAB command line. Make sure you have a file named \textit{input\_array.txt} containing the array elements in \textbf{one line}.
		Or you can just use the UpdateMean, UpdateMedian and UpdateStd files as you wish. Make sure the array A is horizontal.
	}
\end{itemize}
% \\
% \\
\textbf{Solutions: }
% solutions begin here
\begin{enumerate}
	\item{
	Let $\mu$ = mean and $v$ = median of the dataset $\{x_i\}_{i=1}^{N}$ containing $N$ data points.
	$$ \left|\mu - v\right| = \left|\frac{\sum_{i=1}^{N}(x_i - v)}{N}\right| \leq \frac{\sum_{i=1}^{N}\left|x_i - v\right|}{N}$$
	By Triangular Inequality,
	$$ \left|\frac{\sum_{i=1}^{N}x_i - v}{N}\right| \leq \frac{\sum_{i=1}^{N}\left|x_i - v\right|}{N} $$ 
	$$ \implies  \frac{\sum_{i=1}^{N}\left|x_i - v\right|}{N} \leq \frac{\sum_{i=1}^{N}\left|x_i - \mu \right|}{N} $$
	\hfill(the value for which $ \frac{\sum_{i=1}^{N}\left|x_i - x\right|}{N} $  is minimum is when $x$ = median) \\
	Now, \\
	$$ {\Big(\frac{\sum_{i=1}^{N}\left|x_i -\mu\right|}{N}\Big)}^2 \leq \frac{\sum_{i=1}^{N}{\left|x_i -\mu\right|}^2}{N} \leq \frac{\sum_{i=1}^{N}{\left|x_i -\mu\right|}^2}{N-1} = \sigma^2 $$ 
	\hfill (by RMS-AM inequality)
	$$ \therefore \left|\mu - v\right| \leq \sigma $$
	} 


	\item {
	Given 4 datasets $\{x_i\}_{i=1}^{N}$, $\{y_i\}_{i=1}^{N}$ , $\{z_i\}_{i=1}^{N}$, and $\{w_i\}_{i=1}^{N}$ such that,
	$$ z_i = ax_i+b $$ $$ w_i = cy_i+d $$ 
	Using the definition of correlation coefficient, we have,
	$$ r(x,y) = \frac{\sum_{i=1}^{N}{(x_i-\mu_x)(y_i-\mu_y)}}{(N-1)\sigma_x\sigma_y} $$
	$$ r(z,w) = \frac{\sum_{i=1}^{N}{(z_i-\mu_z)(w_i-\mu_w)}}{(N-1)\sigma_z \cdot \sigma_w} $$
	\text{Now if mean of } $\{x_i\}_{i=1}^{N} = \mu_x $\text{ then mean of } $\{z_i\}_{i=1}^{N} = a\mu_x + b$ \text{ , and similarly mean of } $\{w_i\}_{i=1}^{N} = a\mu_y + b$.\\

	Substituting the values of $z_i,\mu_z,w_i,\mu_w$ in the second equation, we get, 
	$$
	r(z,w) = \frac{\sum_{i=1}^{N}{[(ax_i+b)-(a\mu_x+b)] \cdot [(cy_i+d)-(c\mu_y+d)]}}{(N-1)\left|a\right|\sigma_x \cdot \left|c\right|\sigma_y} 
	$$
	$$ =  \frac{\sum_{i=1}^{N}{a(x_i-\mu_x) \cdot c(y_i-\mu_y)}}{(N-1)\left|a\right|\sigma_x \cdot \left|c\right|\sigma_y}$$
	$$ =  sgn(ac)\cdot\frac{\sum_{i=1}^{N}{(x_i-\mu_x)(y_i-\mu_y)}}{(N-1)\sigma_x\sigma_y} $$
	Where $sgn(ac)$ denotes the signum function. \\
	Now, substituting the definition of $r(x,y)$, we have, \\
	$$ r(z,w) = sgn(ac) \cdot r(x,y) $$
	When the sign of $ac$ is positive, i.e. when both $a$ and $c$ are positive or negative, we have $sgn(ac) = 1$, and hence $r(z,w) = r(x,y)$. \\
	When the sign of $ac$ is negative, i.e. when the sign of $a$ and $c$ are opposite, we have $sgn(ac) = -1$, and hence $r(z,w) = -r(x,y)$. \\
	}

	\item {
	Let $x_i$ be the datapoints of a dataset $\{x_i\}_{i=1}^{N}$ with mean $\mu$. \\
	So, we have the following inequality,
	$$	\left|x_i - \mu\right| \leq \sqrt{\sum_{i=1}^{N}{\left|x_i -\mu\right|}^2} $$
		\text{This can easily be shown as follows: }
	$$	\left|x_i -\mu\right|^2 \leq \sum_{i=1}^{N}{\left|x_i -\mu\right|}^2 $$ \\
		\text{Since both R.H.S and L.H.S are positive, taking square root on both sides gives the required result.}
	$$	\left|x_i - \mu\right| \leq \sigma\sqrt{n-1} $$ 
		\hfill \text{ Q.E.D.}
	}

	\item {

	}

	\item {
	Let $ \bar{a} $ denote the mean of a dataset $\{a_i\}_{i=1}^{N}$ having $N$ datapoints. \\ 
	Now, we have, \\
	$$ \bar{a} = \frac{\sum_{i=1}^N{a_i}}{N} $$ \\
	After we add a new element, say $z$, the new mean becomes, \\
	$$\bar{a}_{new} = \frac{(\sum_{i=1}^{N}{a_i} )+ z}{N+1} = \frac{N\bar{a}+z}{N+1} $$
	$\therefore$ We have the value of $\bar{a}_{new}$ in terms of $\bar{a},N,z$. \\ \\

	For the given data, let the standard deviation be denoted by $\sigma$. So we have, \\
	$$\sigma^2 = \frac{\sum_{i=1}^{N}{(x_i-\bar{x})^2}}{N-1} = \frac{(\sum_{i=1}^{N}{x_i^2}) - N\bar{x}^2}{N-1}$$
	
	After adding a new value $z$ to the data, we have a new value $\sigma_{new}$ denoted by, \\
	$$ \sigma_{new}^2 = \frac{\sum_{i=1}^{N+1}{(x_i-\bar{x})^2}}{N} = \frac{(\sum_{i=1}^{N+1}{x_i^2}) - (N+1)\bar{x}_{new}^2}{N} $$
	
	Using the value of $\sigma^2$, we apply the following substitution, \\
	$$ \sum_{i=1}^{N}{x_i^2} = (N-1)\sigma^2 + N\bar{x}^2 $$

	We get the following equation, \\
	$$ \sigma_{new}^2  = \frac{(N-1)\sigma^2 + N\bar{x}^2 + z^2 - (N+1)\bar{x}_{new}^2}{N} $$

	$\therefore$ We have the value of $\sigma_{new}$ in terms of $\sigma$, $\bar{x}$, $\bar{x}_{new}, z$, and $N$. \\
	}
\end{enumerate}

\end{document}